### 2022-04-20 response_reviewers.tex

\section*{Note from the authors}

Dear Dr. Markel,

Thank you for considering our article for publication. We would like to thank the reviewers for their detailed feedback and thoughtful comments and suggestions.

Please find the detailed responses to all the reviewers' comments below. We added line numbers and quotes and highlighted our changes in the manuscript. 
Comments from the Reviewers led to editing and paraphrasing of several parts of the manuscript, including
i) direct mentions of the involvement of the community when defining diversity and inclusion priorities (in Rule 1),
ii) ways to ensure accountability during the organizing process and after the conference (e.g.\ in Rule 2 with a Diversity, Equity, and Inclusion (DEI) team, Rule 8 with guidelines like the diversity statement and the code of conduct, and Rule 10 with post-conference continuity),
iii) mention of women in science in Rule 2,
iv) biases due to video abstracts, 
v) the connection between the code of conduct and the diversity statement, and
v) the process we followed to put together this list of simple rules---now described in detail below and summarized in the current manuscript version. We also included the sets of rules gathered by the team as Supplementary Information.

@@

(1) In Rule 4, the manuscript states `Define measurable indicators for these goals e.g.\ a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions', but how do organizers choose the correct demographics to monitor and the expected background rates? Should it match global, the local, or field-specific rates? Or should it be remediative, where it specifically selects for communities historically ignored in the field beyond the regional population rate?

\begin{Reply}

    % RJ: This is actually rule 1.
   The reviewer asked very important questions, for which there are no straightforward answers or formulae. 
   The `correct' demographic will vary for every conference, field, and local/regional contexts. 
   We believe that while remediation measures should be implemented in all conferences, every meeting committee or organizing team will have to ask themselves these questions and formulate their own inclusive plan.
   We now explicitly mention the need for this reflection and suggest involving the community as well as using available information on diversity in the field (lines \ref{indicators-1} to \ref{indicators-2}):

   \textit{
    Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
    Assess how it translates to the conference's particular field and scientific or professional community.
    Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
    Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.
    }

@@

\end{Reply}

(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants?

\begin{Reply}

The evaluation of goals and outcomes is part of Rule 10, and has now been addressed in the manuscript in lines \ref{info} to \ref{info2}:
\textit{
(...) The organizers will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
}
We also added in Rule 1 an explicit link to Rule 10 in line \ref{rule10} to let the reader know that the evaluation results will be used as part of the conference wrap-up and continuity.

@@

\end{Reply}

If so, should these details be published to the participants after the conference?

\begin{Reply}

Yes, the organizing team should share the final assessment of the success of their inclusive actions with everyone who attended, helped create these goals, or could be part of future conference committees. 
We added more details about this in Rule 10, lines \ref{data-share} to \ref{data-share2}:
\textit{The feedback received from the participants and the whole assessment of the conference should be documented and shared with those who helped define the inclusion goals, the attendees, and future organizers. 
Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents.}
\end{Reply}

@@

How do organizers hold themselves accountable? And to whom are they held accountable to; the organizing committee themselves or the greater community?

\begin{Reply}

Conferences are organized for the community; thus, the organizers should hold themselves accountable to the whole community and not just the organizing committee themselves.
As we mentioned in the manuscript, this should be done along the way (line \ref{rule10}), during the organization process, by being transparent and communicating inclusion efforts and limitations (lines \ref{accountability-1} to \ref{accountability-2}):
\textit{Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.
Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues. 
For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.}

@@

As stated in Rule 10 (lines \ref{account-3} to \ref{data-share2}), there should be a final assessment of the achievement of the diversity and inclusion goals at the end of the conference. Many tools can be used for that (e.g.\ demographic information, surveys, focus groups, information from informal conversations). 
Publishing these final assessments is an important way to hold the organizers accountable both to the greater community and the committee itself.
\end{Reply}

(3) In Rule 4, it states `Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;' Does this pertain to invited speakers only, or speakers selected from submitted abstracts? If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way? Or maybe I'm not fully grasping what curate means in this context?

\begin{Reply}

This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in line \ref{invited}.
We agree with the reviewer that `curation' was an ambiguous term.
We rephrased the section as follows: \textit{
Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} (lines \ref{criteria-1}-\ref{criteria-2}).

@@

I have some comments for authors to consider, and possibly expand upon.

1. As much as I enjoyed reading the ten rules, I missed finding mention of how this is to be implemented and governed. Therefore, it is poignant that authors should brief discuss having an inclusion oversight or governance committee that tracks these metrics and their implementation.

\begin{Reply}

    A Diversity, Equity, and Inclusion (DEI) committee would be best suited to ensure that the inclusion vision and goals are achieved. 
    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}):
    \textit{ 
    Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
    The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} 
    Thank you for pointing this out. 
    On the other hand, the existence of a DEI team is neither necessary nor sufficient, and we believe the most important thing is the commitment of the 
    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2}) and Rule 10 (after the conference, lines \ref{account-3} to \ref{data-share2}) .

@@

2. Secondly, I also missed detailed methodology on how these ten rules were developed or consensus were reached at? Were stakeholders from diverse backgrounds consulted? Was the consultation process iterative? Were the recommendations thus made by those from diverse backgrounds incorporated? How were the final draft reached at - through consensus or voting or a process to that effect? This is important as for such a set of indications to be indicative, it is important that opinions from all backgrounds are taken into account and co-developed.

\begin{Reply}

    The rules were first drafted by the first author, based on the collective experience of organizing useR! 2021 and discussions with other organizers; then, the core group and the members of the diversity, accessibility, and inclusion team were invited to write their own ten simple rules—the rules proposed at that stage are now included in the Supplementary Information.
    We reached a final list by consensus, not popularity. All the other members of the organizing team and the people who helped during the conference (designated helpers) were invited to give their feedback to the resulting list to check for omissions and errors and suggest editions to the text.
    We also sent the manuscript to friendly reviewers who have the experience of attending or organizing conferences. 
    These rules, or their preliminary versions, were shared with 54 people; we received feedback from $\sim$40\%.
    The rules in this manuscript result from this iterative process, discussions with and contributions from all the coauthors, fueled by personal experiences, and extensive literature review.
    A brief description of this process is now stated in lines \ref{methods-a}-\ref{methods-b} of the current version of the manuscript.
    We kept the methodology as brief as possible to focus on the rules instead of our team or conference.

    The coauthors are from diverse backgrounds, career stages, and different continents. They have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
    The manuscript summarizes our work together and is strongly supported by the awareness and experience of organizing a global conference that aims to be a kind, inclusive, and accessible experience for as many people as possible.
\end{Reply}

### 2022-04-20 plos_latex_template.tex
manuscript and responses match

\linelabel{indicators-1}\hl{
Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
Assess how it translates to the conference's particular field and scientific or professional community.
Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering and mathematics) if they exist. 
Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.} 
\linelabel{indicators-2}
Define measurable outcomes for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions). 
These goals and measures will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias} 
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}

@@

\linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
\hl{These documents will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues. 
For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.
\linelabel{accountability-2}

@@

% Rocío: main idea: assess how inclusive was the conference
\linelabel{account-3}As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
\linelabel{info}It is likely that the organizers will be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin}, or preferred language).
The planning committee may include goals concerning how welcomed and included participants felt during the conference, and consider if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees. \linelabel{info2}

% Pass the information
The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers. 
\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data for categories when there is a low number of respondents}.\linelabel{data-share2}
Gather and do your best to organize all the documentation produced during the organization process---or even after. Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors. 
Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
For conference series, it could be a good idea to create an official document (e.g. a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit about the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organizers and defining the selection criteria.

>>>>

\linelabel{indicators-1}\hl{
Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
Assess how it translates to the conference's particular field and scientific or professional community.
Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.} 
\linelabel{indicators-2}
Define measurable outcomes for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions). 
They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias} 
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}

@@

\linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
\hl{These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues. 
For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.
\linelabel{accountability-2}

@@

% Rocío: main idea: assess how inclusive was the conference
\linelabel{account-3}As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
\linelabel{info}\hl{The organizers will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees. \linelabel{info2}

% Pass the information
The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers. 
\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
Gather and do your best to organize all the documentation produced during the organization process---or even after. Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors. 
Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
For conference series, it could be a good idea to create an official document (e.g. a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit about the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organizers and defining the selection criteria.
